---
title: "PCOSP: Pancreatic Cancer Overall Survival Predictor"
author:
    - 'Vandana Sandhu'
    - 'Heewon Seo'
    - 'Christopher Eeles'
    - 'Benjamin Haibe-Kains'
date: 2021-02-01
output:
    BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{PCOSP: Pancreatic Cancer Overall Survival Predictor}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

# Introduction

The PDATK R package provides a set of classes and methods for conducting
survival analysis using biomolecular markers to compute risk scores for patients.
This vignette focuses on implementing

# Classes

## SurvivalExperiment

A `SurvivalExperiment` is a soft wrapper around a `SummarizedExperiment` object
which requires two mandatory metadata columns in the `colData` slot. The
`days_survived` column specifies the integer number of days a patient has
survived since surgery. This is often referred to as the Overall Survival (OS)
of a patient in survival analysis. The `is_deceased` column indicates whether
the patient passed away during the study measurement period and corresponds to
the OS Status of a patient the survival literature. For patients with an
`is_deceased` value of zero survived past the last measurement in the study.

### Constructor

To create a `SurvivalExperiment` object.

```{r SurvivalExperiment_constructor}
library(PDATK)
```



### Accessors

Since a `SurvivalExperiment` contains a `SummarizedExperiment`, all of the
accessor methods are inherited. For more details please see the
`SummarizedExperiment` vignette.

## SurivivalModel

A `SurvivalModel` object.

### Constructor

```{r }`

### Accessors


### Methods


### Sub-Classes

#### PCOSP


#### RandomLabelShufflingModel and RandomGeneAssignmentModel


#### ClinicalModel


#### GeneFuModel

# Pancreate Cancer Overall Survivor Predictor

As an example of the utility of the PDATK package, we will provide code
replicating the analysis published in <inset PCSOP citation>. While the
code presented here is run on dummy data to ensure that PDATK is not too large,
the full data from that study can be downloaded from <where to download it?>.

```{r load_dependencies, message=FALSE, warning=FALSE, include=FALSE}
library(msigdbr)
```
```{r}
data(sampleCohortList)
```

```{r}

```