---
title: "PCOSP: Pancreatic Cancer Overall Survival Predictor"
author:
- name: Vandana Sandhu
- name: Heewon Seo
- name: Christopher Eeles
  affiliation:
  - &pm Bioinformatics and Computational Genomics Laboratory, Princess Margaret Cancer Center,University Health Network, Toronto, Ontario, Canada
  email: christopher.eeles@uhnresearch.ca
- name: Benjamin Haibe-Kains
  affiliation:
  - *pm
  - &mbp Department of Medical Biophysics, University of Toronto, Toronto, Canada
  email: benjamin.haibe.kains@utoronto.ca
date: 2021-02-01
output:
    BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{PCOSP: Pancreatic Cancer Overall Survival Predictor}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

# Pancreate Cancer Overall Survival Predictor

As an example of the utility of the PDATK package, we will provide code
replicating the analysis published in <insert PCSOP citation>. While the
code presented here is run on dummy data to ensure that PDATK is not too large,
the full data from that study can be downloaded from <where to download it?>.

```{r load_dependencies, message=FALSE, warning=FALSE, include=FALSE}
library(PDATK)
library(msigdbr)
```
```{r load_sample_data}
data(sampleCohortList)
sampleCohortList
```

To get started using PDATK, we recommend you place your patient cohorts into
`SurvivalExperiment` objects and then assemble those into a master `CohortList`,
which holds the training and validatino data for use with the various
`SurvivalModel`s in this package.


```{r subset_and_split}
commonGenes <- findCommonGenes(sampleCohortList)
# Subsets all list items, with subset for specifying rows and select for
# specifying columns
cohortList <- subset(sampleCohortList, subset=commonGenes)

ICGCcohortList <- cohortList[grepl('ICGC', names(cohortList), ignore.case=TRUE)]

```