---
title: "Run CSPSC"
author: "Christopher Eeles"
date: "11/05/2020"
output: html_document
---

```{r}
#library(PanCuRx)
require(IDPmisc)
library(dplyr)
library(matrixStats)
library(factoextra)
library(NbClust)
library(cluster)
library(ConsensusClusterPlus)
library(clusterRepro)
library(igraph)
library(gtools)
library(plyr)
library(survival)
library(KMsurv)
library(survminer)
library(limma)
library(piano)
require(ggplot2)
library(ggpubr)
library(BiocParallel)
library(data.table)
library(effsize)
library(vcdExtra)
library(survcomp)
library(RColorBrewer)
library(lattice)
```

## 1. COINCIDE Implementation

```{r load_data}
cohorts <- readRDS('../data/cohortsCommonGenes.rds')       
normalGTEx <- readRDS('../data/normalGTEx.rds')
adjNormal <- readRDS('../data/adjacentNormal.rds')
```


```{r source_function}
source('../R/utilities.R')
```

```{r rorder_cohorts_to_match}
cohortOrder <-c("pcsi", "tcga", 
                "kirby", "icgc_seq", 
                "ouh", "winter", 
                "collisson", "zhang", 
                "grutzmann", "chen", 
                "unc", "icgc_arr", 
                "balagurunathan", "badea", 
                "pei", "haider", 
                "lunardi", "yang", 
                "hamidi", "janky", "bauer")
cohorts <- cohorts[cohortOrder]
```

```{r ranking_genes_based_on_mad}
cohortMADrankings <- bplapply(cohorts, function(cohort) calcRowMADdf(cohort))
```

```{r rank_cohort_genes_by_MAD}
geneWiseWeightedMADdf <- calcGeneWiseWeightedMADdf(cohortMADrankings)
```


```{r get_top_ranked_genes}
top20genesPerCohort <- getTopGenes(cohortMADrankings, n=20)
```


```{r get_top_genes_and_meta_genes}
metaGenes <- getMetaGenes(cohortMADrankings, geneWiseWeightedMADdf, n=20, m=2066)

resultsDir <- file.path("..", "results")
saveRDS(metaGenes, file.path(resultsDir, "metaGenes.rds"))
```



```{r preprocess_cohort_expression_matrixes}
preprocCohorts <- preprocessCohorts(cohorts, metaGenes)
```



```{r compute_consensus_cluster}
consensusClusters <- bplapply(preprocCohorts,
                              consensusClusterCohort,
                              maxK=5, distance="pearson", 
                              method="hc", seed=1987)
```

```{r consensus_cluster_normals_and_GTEx_normals}
normalCohorts <- list("normals"=adjNormal, 
                      "normalsGTEx"=normalGTEx)

preprocNormals <- preprocessCohorts(normalCohorts, metaGenes)
conClustNormals <- bplapply(preprocNormals,
                            consensusClusterCohort,
                            maxK=5, distance="pearson", 
                            method="hc", seed=1987)
```

```{r append_normals_to_cohorts}
allConClusters <- c(consensusClusters, conClustNormals)
allProcCohorts <- c(preprocCohorts, preprocNormals)
```


```{r get_all_pairs}
cohortPairs <- findAllCohortPairs(names(allConClusters))
```


```{r calculate_thresholds}
# 
MSMthresholds <- calcAllMSMthresholds(cohortPairs, allConClusters, allProcCohorts, nthread=16)
```


```{r compare_all_clusters}
a <- Sys.time()
allClusterRepro <- calcAllClusterRepro(MSMthresholds, allConClusters, 
                                       allProcCohorts, nthread=16, numReps=10,
                                       seed=1987)
b <- Sys.time()
b - a
```


```{r }
qsave(allClusterRepro, file.path('..', 'data', 'allClusterRepro.qs'))
```

```{r }

```


## 2. Network Community Search

```{r get_all_cluster_pairs}


```

## 3. Gene Signature Comaprison

```{r }
```

## 4. Biomarker Comparison

```{r }
```

## 5. Single Sample Classifier 

```{r }
```

## 6. Compass Subtyping Waterfall


```{r }
```

## 7. PharmacoGx Cellline Subtyping

```{r }
```

## 8. PharmacoGx Cellline Drug Sensitivity 


```{r }
```

## 9. PDXE Drug Subtyping


```{r }
```


## 10. Cellarity Comparison


```{r }
```

## 11. Copy Number Abberation

```{r }
```

## 12. Caclulate Cluster Meta-Effect Size

```{r }
```

## 13. Pathway Analysis

```{r }
```

## 14. Published Classifier Subtyping

```{r }
```



```{r }
```



