---
title: "PCOSP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Run the script

```{r build_and_select_models}
library(PCOSP)
```

## 1. PCOSP Model Building

```{r load_data}
load(file.path("..", "data", "trainingCohorts.rda"))
```

```{r train_model}
# Initialize a directory for result data
resultsDir <- file.path("..", "results")

# Set seed for calculation (using old R < 3.6 RNG)
set.seed(1987, sample.kind="Rounding")

# Build the models using the selected random seed
system.time({
selectedModels <- buildPCOSPmodels(trainingCohorts, numModels=1000, nthread=15)
})

saveRDS(selectedModels, file=file.path(resultsDir, "1_selectedModels.rds"))
```

## 2. Calculate and Validate Meta-Estimate Statistics

```{r calculate_stats}
load(file.path("..", "data", "validationCohorts.rda"))

# Keep a copy of all the cohorts for later
allValidationCohorts <- validationCohorts

# Select the validation cohorts we want to include
include <- c("TCGA", "PCSI", "Kirby", "ICGC_arr", "UNC", "Chen", "OUH",
             "Zhang", "Winter", "Collisson")

# Subset to selected cohorts
validationCohorts <- validationCohorts[include]

# Fix names to look nice
names(validationCohorts) <- gsub("_arr", "-array", names(validationCohorts))

# Identify which cohorts contain sequencing data
seqCohorts <- c("PCSI", "TCGA", "Kirby")
```


```{r calculate_validation_stats}
validationStats <- calculateValidationStats(validationCohorts,
                                            selectedModels,
                                            seqCohorts=seqCohorts,
                                            nthread=15)

saveRDS(validationStats, file=file.path(resultsDir, "2a_validationStats.rds"))
```

### forestPlotMetaEstimates

```{r forest_plots}
forestPlotMetaEstimate(validationStats, 
                       stat="dIndex", 
                       isSummary=c(rep(FALSE, 10), rep(TRUE, 3)), 
                       filePath=file.path(resultsDir, "figures"),  
                       fileName="2b_forestPlotDindexPCOSP.pdf")
```


```{r forest_plot_cIndex}
forestPlotMetaEstimate(validationStats,
                       stat="cIndex",
                       isSummary=c(rep(FALSE, 10), rep(TRUE, 3)), 
                       filePath=file.path(resultsDir, "figures"),  
                       fileName="2c_forestPlotCindexPCOSP.pdf")
```

### Calculate AUC Distribution and Plot ROC Curves

```{r}
aucStats <- aucMetaEstimates(validationCohorts, validationStats, 
                            seqCohorts=seqCohorts)
saveRDS(aucStats, file=file.path(resultsDir, "2c_aucStatsPCOSP.rds"))
```

```{r format_val_cohorts}
formattedValCohorts <- formatValidationCohorts(validationCohorts)

saveRDS(formattedValCohorts, file=file.path(resultsDir, "2e_formattedValiCohorts.rds"))
```

```{r get_PCOSP_scores}
PCOSPscores <- validationStats$probabilities
```

```{r plot_PCOSP_roc_curve}
colours <- c("chartreuse3", "magenta", "#fb9a99", "turquoise3", "darkgoldenrod1",
             "wheat4", "green", "red", "cornflowerblue", "mediumorchid2")

## FIXME:: Find out where warnings/messages are coming from and suppress them
plotROCcurves(formattedValCohorts, 
              PCOSPscores, 
              colours, 
              filePath=file.path(resultsDir, "figures"), 
              fileName="2f_PCOSProcCurvesPlot")
```

## 3. Model Random Label Shuffling

```{r 3_model_random_label_shuffling}
set.seed(1987, sample.kind="Rounding")

system.time({
    randomLabelModels <- buildRandomLabelShufflingModel(trainingCohorts, 
                                                    numModels=1000, nthread=15)
})
saveRDS(randomLabelModels, file=file.path(resultsDir, "3a_randomLabelModels.rds"))
```

```{r density_plot}
densityPlotModel(formattedValCohorts, 
                 selectedModels=randomLabelModels, 
                 seqCohorts, 
                 vlines=c(0.69, 0.72, 0.70), 
                 nthread=15,
                 title="Random reshuffling of labels",
                 filePath=file.path(resultsDir, "figures"),
                 fileName="3b_densityPlotRandomLabelShuffling.pdf")
```

## 4. Random Gene Assignment Model

```{r 4_random_gene_assignment_models}
set.seed(1987, sample.kind="Rounding")

## These results will differ from the Code Ocean capsule due to sampling.
## For the original results use: original = TRUE; this will be about 5-10x slower
##     than the new method if multiple CPU cores are available.
system.time({
    randomGeneModels <- buildRandomGeneAssignmentModels(trainingCohorts, 
                                                        numModels=1000,
                                                        nthread=15,
                                                        original=TRUE)
})
saveRDS(randomGeneModels, file=file.path(resultsDir, "4a_randomGeneAssignModel.rds"))
```

```{r random_gene_assignment_density}
densityPlotModel(formattedValCohorts, 
                 selectedModels=randomGeneModels, 
                 seqCohorts, 
                 vlines=c(0.69, 0.72, 0.70), 
                 nthread=15,
                 title="Random gene assignment",
                 filePath=file.path(resultsDir, "figures"),
                 fileName="4b_densityPlotRandomGeneAssignment.pdf")
```

## 5. HyperGSA Pathway Analysis

```{r hyper_gsea_models}

dataDir <- file.path("..", "data")
pathwayDataDir <- file.path(dataDir, "Pathway_db_files")
referenceGenes <- read.table(file.path(pathwayDataDir, "reference_genes.txt"))[, 1]
PCOSPgenes <- read.table(file.path(pathwayDataDir, "pcosp_genes.txt"))[, 1]

GSCfiles <- c("hallmark.h.all.v6.1.symbols.gmt",
              "canonical.c2.cp.v6.1.symbols.gmt",
              "GOmolecular.c5.mf.v6.1.symbols.gmt",
              "GOcellular.c5.cc.v6.1.symbols.gmt")

GSCpaths <- vapply(GSCfiles, 
                   function(file, pathwayDataDir) file.path(pathwayDataDir, file),
                   pathwayDataDir,
                   FUN.VALUE=character(1))

GSAmodels <- lapply(GSCpaths,
                  function(filePath, PCOSPgenes, referenceGenes, adjMethod) 
                      fitGSAtoGeneSetCollection(filePath=filePath, 
                                                 PCOSPgenes=PCOSPgenes, 
                                                 referenceGenes=referenceGenes,
                                                 adjMethod=adjMethod),
                  PCOSPgenes=PCOSPgenes,
                  referenceGenes=referenceGenes,
                  adjMethod="fdr"
                  )
names(GSAmodels) <- gsub("\\..*", "", GSCfiles)
saveRDS(GSAmodels, file=file.path(resultsDir, "5_GSAmodels.rds"))
```

## 6. Clinical Model Comparison

```{r clinical_model_comparison}
load(file.path(dataDir, "clinicalFeatures.rda"))
load(file.path(dataDir, "cohortClasses.rda"))

# Rename clinical features for plots and reorder them
names(clinicalFeatures) <- c("PCSI", "ICGCclinical", "TCGA", "ICGCarray", "OUH")
clinicalFeatures <- clinicalFeatures[c("TCGA", "PCSI", "ICGCarray", "OUH", "ICGCclinical")]

fitClinicalModels <- summarizeClinicalModels(clinicalFeatures, cohorts=c("ICGCclinical"))

ICGCclinicalModel <- fitClinicalModels$ICGCclinical

saveRDS(fitClinicalModels, file=file.path(resultsDir, "6a_fitClinicalModels.rds"))
saveRDS(ICGCclinicalModel, file=file.path(resultsDir, "6b_ICGCclinicalModel.rds"))
```


```{r }
# Rename cohortClasses to match cohots
rnCohortClasses <- cohortClasses
names(rnCohortClasses) <- c("PCSI", "TCGA", "ICGC", "Kirby", "ICGCarray", "UNC", "Chen", "OUH", "Winter", "Zhang", "Collison", "ICGCarrayAll")

cohorts <- c("TCGA", "PCSI", "ICGCarray", "OUH")

modelComparisonStats <- compareClinicalModels(clinicalFeatures, rnCohortClasses,
                                   models="ICGCclinical",
                      cohorts=cohorts)
saveRDS(modelComparisonStats, file=file.path(resultsDir, "6c_clinicalModelComparisonStats.rds"))
```

```{r barplot_model_comparison}
barplotModelComparison(modelComparisonStats, 
                       model="ICGCclinical",
                       names=c("TCGA","PCSI","ICGCarray","OUH"),
                       colours=c("palevioletred1", "darkgrey"),
                       filePath=file.path(resultsDir, "figures"),
                       fileName="6d_modelComparisonBarPlot.pdf")
```

```{r calculate_model_comparison_pvals}
modelComparisonPvals <- metaEstimateComparisonAUCs(modelComparisonStats,
                                                   model="ICGCclinical",
                                                   seqCohorts=c("PCSI", "TCGA"))
modelComparisonPvals
saveRDS(modelComparisonPvals, file=file.path(resultsDir, "6e_modelComparisonPvals.rds"))
```

## 7. Clinical Model Dindex and Cindex

```{r clinical_model_dindex_cindex}

clinicalFeatures <- clinicalFeatures[grepl(paste(cohorts, collapse="|"), names(clinicalFeatures))]
names(clinicalFeatures) <- cohorts

modelProbabilities <- calculateCohortProbabilties(fitClinicalModels, clinicalFeatures)

clinicalModelStats <- calculateModelDandCindex(modelProbabilities, 
                                               clinicalFeatures,
                                               seqCohorts=c("PCSI", "TCGA"),
                                               model="ICGCclinical")
saveRDS(clinicalModelStats, file=file.path(resultsDir, "7a_clinicalModelStats.rds"))
```

```{r model_comparison_forest_plot_dindex, fig.height=11, fig.width=8.5}
forestPlotModelComparison(clinicalModelStats, 
                           stat="dIndex",
                           isSummary=c(rep(FALSE, 4), rep(TRUE, 3)),
                           filePath=file.path(resultsDir, "figures"),
                           fileName="7b_modelComparisonDindexForestPlot.pdf")
```


```{r model_comparison_forest_plot_cindex, fig.height=11, fig.width=8.5}
forestPlotModelComparison(clinicalModelStats, 
                           stat="cIndex", 
                           isSummary=c(rep(FALSE, 4), rep(TRUE, 3)),
                           filePath=file.path(resultsDir, "figures"),
                           fileName="7c_modelComparisonCindexForestPlot.pdf")
```

## 8. Existing Classifier Score Calculations

```{r existing_classifier_forest_plots}
dataDir <- file.path("..", "data")
classifierDir <- file.path(dataDir, "Classifier_Comparison")

# Haider data (provided by author)
load(file.path(dataDir, "haiderSigScores.rda"))

# Harmonize case/format of names to use for subsetting the list
names(haiderSigScores) <- gsub("-.*|_.*", "", tolower(names(haiderSigScores)))
    

# Birnbaum data
fileNames1 <- list.files(file.path(classifierDir, "Birnbaum"))[-1]
geneCoefFile1 <-  list.files(file.path(classifierDir, "Birnbaum"))[1]

birnbaumSigScores <- readGenefuFilesToSigScores(file.path(classifierDir, "Birnbaum"), 
                                     fileNames1, 
                                     geneCoefFile1)
saveRDS(birnbaumSigScores, file=file.path(resultsDir, "8a_birnbaumSigScores.rds"))

# Chen data
fileNames2 <- list.files(file.path(classifierDir, "Chen"))[-1]
geneCoefFile2 <- list.files(file.path(classifierDir, "Chen"))[1]

chenSigScores <- readGenefuFilesToSigScores(file.path(classifierDir, "Chen"), 
                                            fileNames2, 
                                            geneCoefFile2, 
                                            signed=TRUE)

saveRDS(chenSigScores, file=file.path(resultsDir, "8b_chenSigScores.rds"))

# PCOSP data
PCOSPscores <- validationStats$probabilities

# Harmonize case/format of names to use for subsetting the list
names(PCOSPscores) <- gsub("-.*|_.*", "", tolower(names(PCOSPscores)))
saveRDS(PCOSPscores, file=file.path(resultsDir, "8c_PCOSPscores.rds"))
```

## 9. Existing Classifier Forestplots

```{r }
validationCohorts <- lapply(validationCohorts, function(cohort) as.data.frame(cohort, row.names=rownames(cohort)))

# Get cohort survival data
cohortSurvivalData <- lapply(validationCohorts, 
                             function(cohort) {
                                 OS <- as.numeric.factor(cohort$OS)
                                 OS_Status <- as.numeric.factor(cohort$OS_Status)
                                 data.frame("OS"=OS, "OS_Status"=OS_Status,
                                     row.names=rownames(cohort))
                                 })

# Fix TCGA data to match study files?
PCOSPscores$tcga <- PCOSPscores$tcga[cohortSurvivalData$TCGA$OS > 180]
cohortSurvivalData$TCGA <- cohortSurvivalData$TCGA[cohortSurvivalData$TCGA$OS > 180, ]

# Harmonize case/format of names to use for subsetting the list
names(cohortSurvivalData) <- gsub("-.*|_.*", "", tolower(names(cohortSurvivalData)))

cohortSurvivalData <- cohortSurvivalData[names(haiderSigScores)]

# Put our classifier data into a list
metaestimateData <- list("haider"=haiderSigScores, "chen"=chenSigScores, "birnbaum"=birnbaumSigScores, 
                         "pcosp"=PCOSPscores, "survival"=cohortSurvivalData)

# Subset cohorts/samples, reorder to match 
metaestimateData <- subsetSharedCohortsAndSamples(metaestimateData)
saveRDS(metaestimateData, file=file.path(resultDir, "9a_metaestimateDataClassifier.rds"))

seqCohorts <- c("tcga", "pcsi", "kirby")

classifierStats <- lapply(metaestimateData[names(metaestimateData) != "survival"],
                          constructMetaEstimatesDF,
                          cohortData=metaestimateData[["survival"]],
                          seqCohorts=seqCohorts,
                          hetero=rep(TRUE, 3))
saveRDS(classifierStats, file=file.path(resultsDir, "9b_classifierStats.rds"))
```

```{r classifier_forest_plot_c_index, fig.height=11, fig.width=8.5}
forestPlotClassifierModelComparision(classifierStats, "cIndex", 
                        names=c("Haider, 2014", "Chen, 2015", "Birnbaum, 2017", "PCOSP"),
                        filePath=file.path(resultsDir, "figures"),
                        fileName="9c_classifierComparisonForestPlot.pdf")
```

```{r classifier_forest_plot_d_index, fig.height=11, fig.width=8.5}
forestPlotClassifierModelComparision(classifierStats, "dIndex", 
                                names=c("Haider, 2014", "Chen, 2015", "Birnbaum, 2017", "PCOSP"),
                                filePath=file.path(resultsDir, "figures"),
                                fileName="9c_classifierComparisonForestPlot.pdf")
```

## 10. Subtype Specific Meta-Estimates

```{r subtype_statistics}
# Ensure all cohorts are data.frames
allValidationCohorts <- lapply(allValidationCohorts, function(cohort) as.data.frame(cohort, row.names=rownames(cohort)))

# Match names with cohortClasses and reorder cohorts to match between class and cohort data
names(allValidationCohorts) <- tolower(names(allValidationCohorts))
names(allValidationCohorts)[which(names(allValidationCohorts) == "icgc_seq")] <- "icgc"

# Remove ICGC_array_all
allValidationCohorts <- allValidationCohorts[!(names(allValidationCohorts) %in% c("icgc_array_all", "icgc"))]
cohortClasses <- cohortClasses[!(names(cohortClasses) %in% c("icgc_array_all", "icgc"))]

# Reorder validation cohrots to same order as the forest plot names
plotNames <- c("tcga", "pcsi", "kirby", "icgc_arr", "unc", "chen", "ouh", "zhang",  "winter", "collisson")
cohortClasses <- cohortClasses[plotNames]
allValidationCohorts <- allValidationCohorts[plotNames]

saveRDS(allValidationCohorts, file=file.path(resultsDir, "10a_allValidationCohorts.rds"))

# Get cohort survival data
allCohortSurvivalData <- lapply(allValidationCohorts, 
                             function(cohort) {
                                 OS <- as.numeric.factor(cohort$OS)
                                 OS_Status <- as.numeric.factor(cohort$OS_Status)
                                 data.frame("OS"=OS, "OS_Status"=OS_Status,
                                 row.names=rownames(cohort))
                                 })
saveRDS(allCohortSurvivalData, file=file.path(resultsDir, "10b_allCohortSurvivalData.rds"))

# Get probabilties for all samples
allCohortProbabilites <- calculatePCOSPscores(allValidationCohorts, selectedModels, nthread=15)
saveRDS(allCohortProbabilites, file=file.path(resultsDir, "10c_allCohortProbabilities.rds"))

# Set sequencing cohorts
seqCohorts <- c("tcga", "pcsi", "kirby")

# Calculate meta-estimates between subtypes per cohort
cohortSubtypeStats <- calculateCohortSubtypeStats(allCohortProbabilites, allCohortSurvivalData, cohortClasses, seqCohorts=seqCohorts)
saveRDS(cohortSubtypeStats, file=file.path(resultsDir, "10d_cohortSubtypeStats.rds"))
```

```{r forest_plot_subtype_comparison_dindex, fig.height=11, fig.width=8.5}
forestPlotCohortSubtypeComparison(cohortSubtypeStats, "dIndex", 
                                  cohortNames=c(NA, "TCGA (n=146)", "Basal-subtype (n=29)", "Classical-subtype (n=117)",
                                                NA, "PCSI (n=118)", "Basal-subtype (n=25)", "Classical-subtype (n=93)",
                                                NA, "Kirby (n=51)", "Basal-subtype (n=13)", "Classical-subtype (n=38)",
                                                NA, "ICGC-array (n=178)", "Basal-subtype (n=156)", "Classical-subtype (n=22)",
                                                NA, "UNC (n=125)", "Basal-subtype (n=27)", "Classical-subtype (n=98)",
                                                NA, "Chen (n=63)", "Basal-subtype (n=20)", "Classical-subtype (n=43)",
                                                NA, "OUH (n=48)", "Basal-subtype (n=08)", "Classical-subtype (n=40)",
                                                NA, "Zhang (n=42)", "Basal-subtype (n=13)", "Classical-subtype (n=29)",
                                                NA, "Winter (n=30)", "Basal-subtype (n=8)", "Classical-subtype (n=22)",
                                                NA, "Collisson (n=27)", "Basal-subtype (n=1)", "Classical-subtype (n=26)",
                                                NA), 
                                  summaryNames=c(NA,"Sequencing", "Basal-subtype", "Classical-subtype",
                                                 NA, "Microarray", "Basal-subtype", "Classical-subtype",
                                                 NA, "Overall", "Basal-subtype", "Classical-subtype"),
                                  filePath=file.path(resultsDir, "figures"),
                                  fileName="10e_subtypeDindexComparisonForestPlot.pdf"
                                  )
```


```{r forest_plot_subtype_comparison_cindex, fig.height=11, fig.width=8.5}
forestPlotCohortSubtypeComparison(cohortSubtypeStats, "cIndex", 
                                  cohortNames=c(NA, "TCGA (n=146)", "Basal-subtype (n=29)", "Classical-subtype (n=117)",
                                                NA, "PCSI (n=118)", "Basal-subtype (n=25)", "Classical-subtype (n=93)",
                                                NA, "Kirby (n=51)", "Basal-subtype (n=13)", "Classical-subtype (n=38)",
                                                NA, "ICGC-array (n=178)", "Basal-subtype (n=156)", "Classical-subtype (n=22)",
                                                NA, "UNC (n=125)", "Basal-subtype (n=27)", "Classical-subtype (n=98)",
                                                NA, "Chen (n=63)", "Basal-subtype (n=20)", "Classical-subtype (n=43)",
                                                NA, "OUH (n=48)", "Basal-subtype (n=08)", "Classical-subtype (n=40)",
                                                NA, "Zhang (n=42)", "Basal-subtype (n=13)", "Classical-subtype (n=29)",
                                                NA, "Winter (n=30)", "Basal-subtype (n=8)", "Classical-subtype (n=22)",
                                                NA, "Collisson (n=27)", "Basal-subtype (n=1)", "Classical-subtype (n=26)",
                                                NA), 
                                  summaryNames=c(NA,"Sequencing", "Basal-subtype", "Classical-subtype",
                                                 NA, "Microarray", "Basal-subtype", "Classical-subtype",
                                                 NA, "Overall", "Basal-subtype", "Classical-subtype"),
                                  filePath=file.path(resultsDir, "figures"),
                                  fileName="10f_subtypeCindexComparisonForestPlot.pdf"
                                  )
```

































